---
title: "Introduction to multibergm"
author: "Brieuc Lehmann"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
date: "6th April, 2020"
vignette: >
  %\VignetteIndexEntry{multibergm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting multibergm

The **multibergm** package can be installed from github and loaded into R using 
the following commands:

```{r setup, message = FALSE}
devtools::install_github("brieuclehmann/multibergm")
library(multibergm)
```

# Statistical network modelling with ERGMs

The **multibergm** package is a set of tools to fit multiple networks using 
Bayesian exponential random graph models (ERGMs). An ERGM defines a parametric 
statistical distribution across all possible networks with a given set of nodes.
The aim of the model is to characterise the distribution of a network in terms 
of a set of summary statistics. These summary statistics are typically comprised
of topological features of the network, such as the number of edges and subgraph 
counts.

**multibergm** is built on and inspired by the **ergm** R package, and uses the 
same syntax to specify models. For a great introduction to the **ergm** package,
see the following vignette:

```{r ergm-vignette, eval = FALSE}
vignette("ergm")
```

# Loading data

To get started with some network data, the **multibergm** package includes some 
synthetic networks. We'll use this data to demonstrate the functionality of 
**multibergm**. You can load these by entering the following commands.

```{r ergm-vignette, eval = FALSE}
vignette("ergm")
```

The package also provides a utility function to create networks from correlation 
matrices. 

Example data (FIND SOME... OR SIMULATE?)

# The multibergm framework

Describe the hierarchical framework

# Model initialisation

Priors, initial values, hyperparameters etc.

# Model fitting

Main function goes here

# Goodness of fit

Visualisation etc.
